---
title: 
author: 
date: 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css echo=FALSE}
p {
    font-family: "Open Sans";   
}
```
<style>
.html-widget {
    margin: auto;
}
</style>

```{r, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#install packages 
library(sf)
library(dplyr)
library(leaflet)
library(stringi)
library(RColorBrewer)
library(tidyverse)
library(ggplot2)
library(rgdal)
library(kableExtra)
library(scales)
library(ggrepel)
library(formattable)
library(tmap)
library(fontawesome)
library(magick)
library(mapview)
library(extrafont)
library(ceramic)
library(mapboxapi)
library(knitr)
font_import("C:/Users/GraceParker/Downloads/Open_Sans")
# set working directory 

#insert data
data_raw <- read.csv("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/BIL_AwardSummaries_2023-07-03.csv")

territories <- st_read("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/Territories/", quiet=TRUE)

#clean BIL data
# targetD = toupper(targetC)
targetF <- c("01","05","12","13","21","22","28","37","45","47","51","72","60","69","66","78")
targetG <- c(1,5,12,13,21,22,28,37,45,47,51,72,60,69,66,78)
data <-select(data_raw, awarding_agency_name, total_obligated_amount, recipient_state_name, recipient_county_name, prime_award_summary_recipient_state_fips_code)
SETerr <- dplyr::filter(territories,FIPS %in% targetF)
targetH <- c("01","05","12","13","21","22","28","37","45","47","51")
SEstates <- dplyr::filter(territories,FIPS %in% targetH)
SETerrData <- dplyr::filter(data,prime_award_summary_recipient_state_fips_code %in% targetG)

#group funding by territory/state
fundByState <- group_by(SETerrData, prime_award_summary_recipient_state_fips_code) %>% dplyr::summarise(
  StateFunding = sum(total_obligated_amount))%>%
  as.data.frame()

SETerr$FIPS <- as.numeric(SETerr$FIPS)
colnames(fundByState)[1] <- "FIPS"
#join fundbyCounty to SEcounties
BIL <- left_join(SETerr, fundByState, by = "FIPS")

bins <- c(1000000000, 1500000000, 2000000000, 2500000000,3000000000,Inf)
pal <- colorBin(c("#B8E9F0","#75C7D3", "#00869F", "#006A7E","#00447C"), domain = BIL$StateFunding, bins=bins)

state_popup <- paste0("<strong>State Name: </strong>", 
                      str_to_title(BIL$NAME), 
                      "<br><strong>Total BIL Funding: </strong>", 
                      formattable::currency(BIL$StateFunding, digits = 0L))

#insert function/award type and population data
Asst2023_raw <- read.csv("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/BIL_FY2023_P8_wFunction/FY2023P01-P08_Assistance.csv")
Con2023_raw <- read.csv("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/BIL_FY2023_P8_wFunction/FY2023P01-P08_Contracts.csv")
Unl2023_raw <- read.csv("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/BIL_FY2023_P8_wFunction/FY2023P01-P08_Unlinked.csv")

Asst2022_raw <- read.csv("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/BIL_FY2022_wFunction/FY2022_Assistance.csv")
Con2022_raw <- read.csv("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/BIL_FY2022_wFunction/FY2022_Contracts.csv")
Unl2022_raw <- read.csv("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/BIL_FY2022_wFunction/FY2022_Unlinked.csv")

pop2022_raw <- read.csv("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/POP2022.csv")

#append data together
funData <- rbind(Asst2023_raw, Con2023_raw)
funData <- rbind(funData, Unl2023_raw)
funData <- rbind(funData, Asst2022_raw)
funData <- rbind(funData, Con2022_raw)
funData <- rbind(funData, Unl2022_raw)

#select
funData <- funData[!is.na(funData$transaction_obligated_amount),]
funData <-select(funData, budget_subfunction, awarding_agency_name, awarding_subagency_name, transaction_obligated_amount, recipient_state, recipient_county, award_type)

#select population values
colnames(pop2022_raw)[1] <- "State"
pop2022 <-select(pop2022_raw, State, X.3)

#get total and perCap for US
totalUS <- formattable::currency(sum(funData$transaction_obligated_amount), digits=0L)
pop2022$X.3 <- as.numeric(gsub(",","",pop2022$X.3))
totPopUS <- pop2022$X.3[4]
avgPerCapUS <- formattable::currency(totalUS/totPopUS, digits=0L)

#filter
targetE <- c("AL","AR","FL","GA","KY", "LA", "MS", "NC", "SC", "TN", "VA", "GU", "PR", "MP", "AS", "VI")
targetO <- c("GU", "PR", "MP", "AS", "VI")
funData <- dplyr::filter(funData, recipient_state %in% targetE)
DataTerr <- dplyr::filter(funData, recipient_state %in% targetO)
funData$Region <- "Southeast"
sankey_SE_Data <- select(funData, budget_subfunction, transaction_obligated_amount, Region)
sankey_SE_Data2 <- select(funData, award_type, transaction_obligated_amount, Region)
# colnames(sankey_SE_Data2)[1] <- "budget_subfunction"
# colnames(sankey_SE_Data2)[3] <- "award_type"
# sankey_SE_Data <- as.data.frame(rbind(sankey_SE_Data, sankey_SE_Data2))
write.csv(sankey_SE_Data,file="C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/funData_sankey_states.csv")
#filter pop by state
targetF <- c(".Alabama",".Arkansas",".Florida",".Georgia",".Kentucky", ".Louisiana", ".Mississippi", ".North Carolina", ".South Carolina", ".Tennessee", ".Virginia")
pop2022 <- dplyr::filter(pop2022, State %in% targetF)

#group by state and sum funding
stateData <- funData %>%                                     
  group_by(recipient_state) %>%
  dplyr::summarise(Funding = sum(transaction_obligated_amount)) %>% 
  as.data.frame()

#calculate percent and per cap funding
colnames(stateData)[1] <- "State"
pop2022$State <- c('AL', 'AR', 'FL', 'GA', 'KY', 'LA', 'MS', 'NC', 'SC', 'TN', 'VA')
stateData <- left_join(stateData, pop2022, by = "State") 
stateData$Funding <- as.numeric(stateData$Funding)
stateData$X.3 <- as.numeric(gsub(",","",stateData$X.3))
totPop <- sum(stateData$X.3)
stateData$perCapita <- stateData$Funding/stateData$X.3
stateData$Percent <- stateData$Funding/sum(stateData$Funding)*100
stateData$Percent <- format(round(stateData$Percent, 2), nsmall=2)
stateData$Percent <- as.numeric(stateData$Percent)
stateData$Funding <- signif(stateData$Funding, 5)
stateData$Funding <- formattable::currency(stateData$Funding, digits = 0L)
stateData$perCapita <- formattable::currency(stateData$perCapita, digits = 0L)
pop2022$X.3 <- as.numeric(gsub(",","",pop2022$X.3))
totPop <- sum(pop2022$X.3)

# #Investigate negative values
# deoblig <- filter(funData, transaction_obligated_amount < 0)
# 
# deoblig1 <- deoblig %>% 
#   group_by(recipient_state) %>% dplyr::summarise(
#   Funding = sum(transaction_obligated_amount)
#   )%>% as.data.frame()
# 
# deTot <- sum(deoblig1$Funding)
#   
# deoblig1$Percent <- 100*(deoblig1$Funding/sum(deoblig1$Funding))
# deoblig1$Percent <- format(round(deoblig1$Percent, 2), nsmall=2)
# deoblig1$Percent <- as.numeric(deoblig1$Percent)
# 
# deoblig1$zone.start <- with(deoblig1, c(0, cumsum(Percent)[-length(Percent)]))  
# deoblig1$zone.end <- with(deoblig1, cumsum(Percent))  
# deoblig1$label.point <- with(deoblig1, (100-(zone.start + zone.end) / 2)) 
# 
# colnames(deoblig1)[1] <- "State"
# 
# pie <- ggplot(deoblig1, aes(x="", y=Percent, fill=State))+
#   geom_bar(width = 1, stat = "identity")+
#   coord_polar("y", start=0)+
#   scale_fill_manual(values=c("#FFA07A","#900C3F", "#C70039", "#D4442D", "#00869F", "#6171AF","#DCB6D0","#C799C8", "#D2D2D2", "#7B7B7B","#585757","#17202A")) +
#   geom_text_repel(data = deoblig1,
#                    aes(y = label.point, label = ifelse(Percent>10,paste0(Percent, "%"),""), family="Open Sans"),
#                    size = 4.5, nudge_x = 1, show.legend = FALSE) +
#   ggtitle("Southeast Deobligated Funding by State")+
#   theme_void(base_family= "Open Sans")+
#   theme(legend.key.height = unit(.5,'cm'),
#         legend.key.width = unit(.5, 'cm'))
# pie
# deoblig1$Funding <- formattable::currency(deoblig1$Funding, digits = 0L)
# deoblig1 <- deoblig1%>%filter(State != "PR")
# deoblig1$State[1] <- "Alabama"
# deoblig1$State[2] <- "Arkansas"
# deoblig1$State[3] <- "Florida"
# deoblig1$State[4] <- "Georgia"
# deoblig1$State[5] <- "Kentucky"
# deoblig1$State[6] <- "Louisiana"
# deoblig1$State[7] <- "Mississippi"
# deoblig1$State[8] <- "North Carolina"
# deoblig1$State[9] <- "South Carolina"
# deoblig1$State[10] <- "Tennessee"
# deoblig1$State[11] <- "Virginia"
# deoblig1$Funding <- abs(deoblig1$Funding)
# deoblig1 <- deoblig1[order(deoblig1$Funding,decreasing=TRUE),]
# DeobTable<- deoblig1%>%select(State, Funding)%>% kbl(row.names=FALSE) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, html_font="Calibri", font_size = 14, position="center", row_label_position = r)%>%
#   add_header_above(c("Deobligated Funding by State"=2))%>%
#   row_spec(1:11, color  = "black")%>%
#   column_spec(1:2, "3cm")
# DeobTable
# 
# save_kable(DeobTable,"C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/Deobtable.png")
# 
# deobligPR <- deoblig %>% filter(recipient_state== "PR")%>%
#   group_by(budget_subfunction) %>% dplyr::summarise(
#   Funding = sum(transaction_obligated_amount)
#   )%>% as.data.frame()
# 
# deobligPR2 <- deoblig %>% filter(recipient_state== "PR")%>%
#   group_by(award_type) %>% dplyr::summarise(
#   Funding = sum(transaction_obligated_amount)
#   )%>% as.data.frame()
# 
# 
# #By Category
# deoblig2 <- deoblig %>% 
#   group_by(budget_subfunction) %>% dplyr::summarise(
#   Funding = sum(transaction_obligated_amount)
#   )%>% as.data.frame()
# 
# deoblig2$Percent <- 100*(deoblig2$Funding/sum(deoblig2$Funding))
# deoblig2$Percent <- format(round(deoblig2$Percent, 2), nsmall=2)
# deoblig2$Percent <- as.numeric(deoblig2$Percent)
# 
# deoblig2$zone.start <- with(deoblig2, c(0, cumsum(Percent)[-length(Percent)]))  
# deoblig2$zone.end <- with(deoblig2, cumsum(Percent))  
# deoblig2$label.point <- with(deoblig2, (100-(zone.start + zone.end) / 2)) 
# 
# colnames(deoblig2)[1] <- "Category"
# 
# pie <- ggplot(deoblig2, aes(x="", y=Percent, fill=Category))+
#   geom_bar(width = 1, stat = "identity")+
#   coord_polar("y", start=0)+
#   scale_fill_manual(values=c("#FFA07A","#900C3F", "#C70039", "#D4442D", "#00869F", "#6171AF","#DCB6D0","#C799C8", "#D2D2D2", "#7B7B7B","#585757","#17202A")) +
#   geom_text_repel(data = deoblig2,
#                    aes(y = label.point, label = ifelse(Percent>5 | Category == 'Energy conservation' ,paste0(Percent, "%"),""), family="Open Sans"),
#                    size = 4.5, nudge_x = 1, show.legend = FALSE) +
#   ggtitle("Southeast Deobligated Funding by Category")+
#   theme_void(base_family= "Open Sans")+
#   theme(legend.key.height = unit(.5,'cm'),
#         legend.key.width = unit(.5, 'cm'))
# pie
# 
# ggsave("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/Categories_Deob_SE.png", width = 6,   height = 4.25, dpi=1100)
# 
# #by Type
# deoblig3 <- deoblig %>% 
#   group_by(award_type) %>% 
#   dplyr::summarise(
#   Funding = sum(transaction_obligated_amount)
#   )%>% as.data.frame()
# 
# deoblig3$Percent <- 100*(deoblig3$Funding/sum(deoblig3$Funding))
# deoblig3$Percent <- format(round(deoblig3$Percent, 2), nsmall=2)
# deoblig3$Percent <- as.numeric(deoblig3$Percent)
# 
# deoblig3$zone.start <- with(deoblig3, c(0, cumsum(Percent)[-length(Percent)]))  
# deoblig3$zone.end <- with(deoblig3, cumsum(Percent))  
# deoblig3$label.point <- with(deoblig3, (100-(zone.start + zone.end) / 2)) 
# 
# colnames(deoblig3)[1] <- "Type"
# deoblig3$Type <- str_to_title(deoblig3$Type)
# deoblig3$Type[6] <- "Other Financial Assistance"
# deoblig3$Type[2] <- "Cooperative Agreement"
# deoblig3$Type[1] <- "BPA Call"
# deoblig3$Type[5] <- "Formula Grant"
# deoblig3$Type[7] <- "Project Grant"
# 
# pie <- ggplot(deoblig3, aes(x="", y=Percent, fill=Type))+
#   geom_bar(width = 1, stat = "identity")+
#   coord_polar("y", start=0)+
#   scale_fill_manual(values=c("#900C3F", "#D4442D", "#FCB616", "#75C7D3", "#00869F", "#C799C8", "#D2D2D2", "#585757")) +
#   geom_text_repel(data = deoblig3,
#                    aes(y = label.point, label = ifelse(Percent>5,paste0(Percent, "%"),""),family="Open Sans"),
#                    size = 4.5, nudge_x = 1, show.legend = FALSE, max.overlaps =20) +
#   ggtitle("Southeast Deobligated Funding by Award Type")+
#   theme_void(base_family= "Open Sans")+ 
#   theme(plot.title = element_text(size=12))
# pie
```

## Bipartisan Infrastructure Law Funding in the Territories

<div class = "row">
<div class = "col-md-8" style="padding-right:5px;">
```{r, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
BIL$StateFunding <- formattable::currency(BIL$StateFunding)

terrList <- c("Puerto Rico", "American Samoa", "Guam", "Northern Mariana Islands", "Virgin Islands (U.S.)")
terrFIPS <- c(60,66,69,72,78)
terr_pop <- read.csv("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/Territories_pop/Territories_pop.csv")%>%
  select(Country.Name, X2022)%>%
  dplyr::filter(Country.Name %in% terrList)
# fundByTerr$Funding <- formattable::currency(fundByTerr$Funding, digits = 0L)
targetN <- c("AS","GU", "MP", "PR", "VI")
terr_pop$State <- c("AS","GU","MP","PR","VI")
terrData <- left_join(stateData, terr_pop, by = "State")%>%
  dplyr::filter(State %in% targetN)
colnames(terrData)[6] <- "Territory"

gTarg <- c("COOPERATIVE AGREEMENT (B)","PROJECT GRANT (B)")
Grants_Terr <- dplyr::filter(funData, award_type %in% gTarg)%>%dplyr::filter(recipient_state %in% targetN)%>%
  group_by(recipient_state)%>%
  dplyr::summarise(Funding = sum(transaction_obligated_amount)) %>% 
  as.data.frame()

terrData$perCapita <- terrData$Funding/terrData$X2022
colnames(Grants_Terr)[1] <- "State"
terrData <- left_join(terrData,Grants_Terr, by="State")
terrData$GrantsPerCap = terrData$Funding.y/terrData$X2022
terrData$GrantsPerCap <- formattable::currency(terrData$GrantsPerCap, digits=0L)
terrData$Funding.y <- signif(as.numeric(terrData$Funding.y), 5)
terrData$Funding.y <- formattable::currency(terrData$Funding.y, digits=0L)
 
terrData<- terrData%>% select(Territory, Funding.x, perCapita, Funding.y, GrantsPerCap) 
terrData <- terrData[order(terrData$Funding.x,decreasing=TRUE),]

# colnames(Grants_Terr)[1] <-"STATE"
TerrMap <- SETerr%>% filter(STATE%in% targetN)
colnames(TerrMap)[2] <- "Territory"
terrData$Territory[5] <- "Virgin Islands"
TerrGrantsMap <- left_join(TerrMap, terrData, by="Territory")
Total <-TerrGrantsMap
Grants <- TerrGrantsMap
#tmap PR
tmap_mode(mode ="view")
puerto <- tm_basemap("CartoDB.PositronNoLabels")+
tm_shape(Total) +
  tm_polygons("Funding.x",
      palette= c("#75C7D3", "#006A7E", "#00315E"),
      breaks=c(0, 50000000, 100000000, Inf),
      textNA= "$0",
      labels = c("$0-50 million", "$50-100 million", "$100 million+"),
      legend.position=c("right","bottom"),
      popup.vars= c("Funding: "="Funding.x"),
      popup.format= list(prefix="$", big.num.abbr=NA),
      id= "NAME",
      title= "Territory Funding",
      border.col="white")+
  tm_shape(Grants) +
  tm_polygons("Funding.y",
      palette= c("#75C7D3", "#006A7E", "#00315E"),
      breaks=c(0, 30000000, 50000000, Inf),
      textNA= "$0",
      labels = c("$0-30 million", "$30-50 million",
      "$50 million+"),
      legend.show=FALSE,
      popup.vars= c("Grants: "="Funding.y"),
      popup.format= list(prefix="$", big.num.abbr=NA),
      id= "STATE",
      title= "Grants Funding",
      border.col="white")+
tm_borders(col="white", lwd=.25)+
tmap_options(check.and.fix = TRUE)+
tm_view(set.view=c(-65.5,18,7))

#tmap GU
tmap_mode(mode ="view")
guam <- tm_basemap("CartoDB.PositronNoLabels")+
tm_shape(Total) +
  tm_polygons("Funding.x",
      palette= c("#75C7D3", "#006A7E", "#00315E"),
      breaks=c(0, 50000000, 100000000, Inf),
      textNA= "$0",
      labels = c("$0-50 million", "$50-100 million", "$100 million+"),
      legend.show=FALSE,
      popup.vars= c("Funding: "="Funding.x"),
      popup.format= list(prefix="$", big.num.abbr=NA),
      id= "NAME",
      title= "Territory Funding",
      border.col="white")+
  tm_shape(Grants) +
  tm_polygons("Funding.y",
      palette= c("#75C7D3", "#006A7E", "#00315E"),
      breaks=c(0, 30000000, 50000000, Inf),
      textNA= "$0",
      labels = c("$0-30 million", "$30-50 million",
      "$50 million+"),
      popup.vars= c("Grants: "="Funding.y"),
      popup.format= list(prefix="$", big.num.abbr=NA),
      id= "STATE",
      title= "Grant Funding",
      border.col="white")+
tm_borders(col="white", lwd=.25)+
tmap_options(check.and.fix = TRUE)+
tm_view(set.view=c(144.5,14,8))

#tmap AS
samoa <- tm_basemap("CartoDB.PositronNoLabels")+
tm_shape(Total) +
  tm_polygons("Funding.x",
      palette= c("#75C7D3", "#006A7E", "#00315E"),
      breaks=c(0, 50000000, 100000000, Inf),
      textNA= "$0",
      labels = c("$0-50,000,000", "$50,000,000-100,000,000", "$100,000,000+"),
      legend.show=FALSE,
      popup.vars= c("Funding: "="Funding.x"),
      popup.format= list(prefix="$", big.num.abbr=NA),
      id= "NAME",
      title= "Territory Funding",
      border.col="white")+
  tm_shape(Grants) +
  tm_polygons("Funding.y",
      palette= c("#75C7D3", "#006A7E", "#00315E"),
      breaks=c(0, 30000000, 50000000, Inf),
      textNA= "$0",
      labels = c("$0-50,000,000", "$50,000,000-100,000,000",
      "$100,000,000+"),
      legend.show=FALSE,
      popup.vars= c("Grants: "="Funding.y"),
      popup.format= list(prefix="$", big.num.abbr=NA),
      id= "STATE",
      title= "Grants Funding",
      border.col="white")+
tm_borders(col="white", lwd=.25)+
tmap_options(check.and.fix = TRUE)+
tm_view(set.view=c(-170,-14,7))

tmap_arrange(puerto, guam, samoa, nrow=1, ncol=3, widths=c(.25,.25,.25))

# BILterr <- BIL%>%filter(STATE %in% targetO)
# chartTerr <- left_join(BILterr,Grants_Terr, by="STATE")
# 
# chartTerr <- chartTerr[order(chartTerr$Funding,decreasing=TRUE),]%>%as.data.frame()%>%
#   select(STATE, Funding, perCapita, GrantFunding, perCapitaGrants)
# colnames(RegMap)[3] <-"Total Per Capita"
# colnames(RegMap)[4] <- "Grant Funding"
# colnames(RegMap)[5] <- "Grants Per Capita"
# RegionTable<- RegMap%>% kbl(row.names=FALSE) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, html_font="Calibri", font_size = 14, position="center", row_label_position = r)%>%
#   add_header_above(c("Funding by Region"=5))%>%
#   row_spec(1:9, color  = "black")%>%
#   column_spec(1:5, "3cm")
# RegionTable
# 
# save_kable(RegionTable,"C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/Regiontable.png")
# 
# bbox_PR <- st_bbox(c(xmin = -67.7, xmax = -64.0, ymax = 19, ymin = 17.5), crs = st_crs(4326))
# 
# mbtiles <- get_static_tiles(
#   location = c(-67.55, 17.3, -63.85, 19.1),
#   zoom = 7,
#   style_url= "mapbox://styles/gparker-seea/clkr4hgj800zo01qk7npq3p0x",
#   username="gparker-seea",
#   access_token= "pk.eyJ1IjoiZ3Bhcmtlci1zZWVhIiwiYSI6ImNsa3I0ZHFtajJnc2QzcWtlajlmbWV3d2IifQ.jk6o3SXtTTAneVg11e_pww"
# )
# 
# tm_shape(mbtiles) +
#   tm_rgb() +
#   tm_shape(TerrGrantsMap) +
#   tm_polygons("Funding",
#           palette= c("#75C7D3", "#006A7E", "#00315E"),
#           breaks=c(0, 30000000, 50000000, Inf),
#           labels = c("$0-30 million", "$30-50 million", "$50 million+"),
#           textNA= "$0",
#           alpha=.85,
#       legend.show=FALSE,
#       border.col="white") +
#   tm_layout(legend.outside = TRUE)
# 
# tmap_save(filename="C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/puerto_grants.png", dpi=1200)
# 
# bbox_AS <- st_bbox(c(xmin = -171.2, xmax = -169.0, ymax = -13.9, ymin = -14.5), crs = st_crs(4326))
# 
# mbtiles <- get_static_tiles(
#   location = c(-171, -13.9, -169.3, -14.65),
#   zoom = 8,
#   style_url= "mapbox://styles/gparker-seea/clkr4hgj800zo01qk7npq3p0x",
#   username="gparker-seea",
#   access_token= "pk.eyJ1IjoiZ3Bhcmtlci1zZWVhIiwiYSI6ImNsa3I0ZHFtajJnc2QzcWtlajlmbWV3d2IifQ.jk6o3SXtTTAneVg11e_pww"
# )
# 
# # tm_shape(mbtiles) +
# #   tm_rgb() +
# #   tm_shape(TerrGrantsMap) +
# #   tm_polygons("Funding",
# #           palette= c("#75C7D3", "#006A7E", "#00315E"),
# #           breaks=c(0, 30000000, 50000000, Inf),
# #           textNA= "$0",
# #           alpha=.85,
# #       legend.show=FALSE,
# #       border.col="white")+
# #   tm_layout(legend.outside = TRUE)
# 
# tmap_save(filename="C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/samoa_grants.png", dpi=1200)


# bbox_GU <- st_bbox(c(xmin = 144, xmax = 146.7, ymax = 15.4, ymin = 13), crs = st_crs(4326))
# mbtiles <- get_static_tiles(
#   location = c(145.05, 13.9, 145.85, 15.5),
#   zoom = 7,
#   style_url= "mapbox://styles/gparker-seea/clkr4hgj800zo01qk7npq3p0x",
#   username="gparker-seea",
#   access_token= "pk.eyJ1IjoiZ3Bhcmtlci1zZWVhIiwiYSI6ImNsa3I0ZHFtajJnc2QzcWtlajlmbWV3d2IifQ.jk6o3SXtTTAneVg11e_pww"
# )
# 
# tm_shape(mbtiles) +
#   tm_rgb() +
#   tm_shape(BIL) +
#   tm_polygons("StateFunding",
#           palette= c("#75C7D3", "#006A7E", "#00315E"),
#           breaks=c(0, 50000000, 100000000, Inf),
#           labels = c("$0-50,000,000", "$50-100 million", "$100 million+"),
#           textNA= "$0",
#           alpha=.85,
#       legend.show=FALSE,
#       border.col="white")
# 
# # tmap_save(filename="C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/mariana.png", dpi=1200)

# mbtiles <- get_static_tiles(
#   location = c(144.4, 12.5, 145.2, 14.1),
#   zoom = 7,
#   style_url= "mapbox://styles/gparker-seea/clkr4hgj800zo01qk7npq3p0x",
#   username="gparker-seea",
#   access_token= "pk.eyJ1IjoiZ3Bhcmtlci1zZWVhIiwiYSI6ImNsa3I0ZHFtajJnc2QzcWtlajlmbWV3d2IifQ.jk6o3SXtTTAneVg11e_pww"
# )
# 
# tm_shape(mbtiles) +
#   tm_rgb() +
#   tm_shape(BIL) +
#   tm_polygons("StateFunding",
#           palette= c("#75C7D3", "#00869F", "#00315E"),
#           breaks=c(0, 50000000, 100000000, Inf),
#           labels = c("$0-50 million", "$50-100 million", "$100 million+"),
#           textNA= "$0",
#           alpha=.85,
#           title= "BIL Territory Funding",
#       # legend.show=FALSE,
#       border.col="white") +
#   tm_layout(legend.outside = TRUE, legend.outside.position="bottom", legend.text.fontfamily = "Calibri",
#       legend.title.fontfamily = "Calibri",
#       legend.title.size= 1,
#       legend.text.size=1)

# tmap_save(filename="C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/guam.png", dpi=1200)

```
</div>

<div class = "col-md-4">
<br>
The Infrastructure Investment and Jobs Act, also known as the Bipartisan Infrastructure Law (BIL), was passed in November 2021 and allocates $1.2 trillion over ten years in funding for infrastructure projects including roads, public transit, broadband, and electric grid upgrades. As federal agencies continue to allocate BIL funding, this map shows how much funding recipients in the territories have received as of July 13, 2023. This dashboard focuses on funding available through the BIL, and we plan to address additional sources of funding, including the Inflation Reduction Act (IRA) in future versions of this whitepaper as more funding is allocated.
<br>
</div>
## Funding Analysis {.tabset}

### Territories
<div class = "row">
<div class = "col-md-7" style="padding-right:25px;">
``` {r echo=FALSE}
terrData$Percent <- 100*(as.numeric(terrData$Funding.x)/sum(as.numeric(terrData$Funding.x)))
terrData$Percent <- format(round(terrData$Percent, 2), nsmall=2)
terrData$Percent <- as.numeric(terrData$Percent)
terrData$zone.start <- with(terrData, c(0, cumsum(Percent)[-length(Percent)]))  
terrData$zone.end <- with(terrData, cumsum(Percent))  
terrData$label.point <- with(terrData, ((zone.start + zone.end) / 2)) 

pie <- ggplot(terrData, aes(x="", y=Percent, fill=reorder(Territory, Percent, decreasing=F)))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=c("#900C3F", "#C70039", "#D4442D", "#FCB616","#FFF7B3"), name="Territories") +
  geom_text_repel(data = terrData,
                   aes(y = label.point, label = ifelse(Percent>5,paste0(Percent, "%"),""), family="Open Sans"),
                   size = 4.5, nudge_x = 1, show.legend = FALSE) +
  ggtitle("U.S. Territories Funding")+
  theme_void(base_family= "Open Sans")+
  theme(legend.key.height = unit(.5,'cm'),
        legend.key.width = unit(.5, 'cm'))
pie

ggsave("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/FundingbyTerr.png", width = 6,   height = 4.25, dpi=1100)
```
</div>

<div class = "col-md-5">
``` {r echo=FALSE}

terrData <-terrData%>%select(Territory, Funding.x, perCapita)
colnames(terrData)[2] <- "Funding" 
colnames(terrData)[3] <- "Total Per Capita" 
# colnames(terrData)[4] <- "Grant Funding" 
# colnames(terrData)[5] <- "Grants Per Capita" 
terrTable<- terrData%>% kbl(row.names=FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, html_font="Open Sans", font_size = 14, position="center", row_label_position = r)%>%
  add_header_above(c("Funding by Territory"=3))%>%
  row_spec(1:3, color  = "black")%>%
  column_spec(1:3, "4cm")
terrTable
save_kable(terrTable,"C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/Terrtable.png")


terrTotal <- sum(terrData$Funding)
percTerr <- as.numeric(terrTotal/totalUS*100)
terrPoptotal <- sum(terrData$X2022)
perCapitaTerr <- terrTotal/terrPoptotal

```
</div>
### Funding by Category
<div class = "row">
<div class = "col-md-7" style="padding-right:25px;">
``` {r echo=FALSE, fig.align='center'}
funcTerr <- funData %>% dplyr::filter(recipient_state %in% targetN)%>%
  group_by(budget_subfunction) %>% dplyr::summarise(
  Funding = sum(transaction_obligated_amount)
  )%>% as.data.frame()

funcTerr$Percent <- 100*(funcTerr$Funding/sum(funcTerr$Funding))
funcTerr$Percent <- format(round(funcTerr$Percent, 2), nsmall=2)
funcTerr$Percent <- as.numeric(funcTerr$Percent)

funcTerr$zone.start <- with(funcTerr, c(0, cumsum(Percent)[-length(Percent)]))  
funcTerr$zone.end <- with(funcTerr, cumsum(Percent))  
funcTerr$label.point <- with(funcTerr, (100-(zone.start + zone.end) / 2)) 

colnames(funcTerr)[1] <- "Category"

pie <- ggplot(funcTerr, aes(x="", y=Percent, fill=Category))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=c("#FFA07A","#900C3F", "#C70039", "#D4442D", "#00869F", "#6171AF","#DCB6D0","#C799C8", "#D2D2D2", "#7B7B7B","#585757","#17202A")) +
  geom_text_repel(data = funcTerr,
                   aes(y = label.point, label = ifelse(Percent>5 | Category == 'Energy conservation' ,paste0(Percent, "%"),""), family="Open Sans"),
                   size = 4.5, nudge_x = 1, show.legend = FALSE) +
  ggtitle("U.S. Territories Funding by Category")+
  theme_void(base_family= "Open Sans")+
  theme(legend.key.height = unit(.5,'cm'),
        legend.key.width = unit(.5, 'cm'))
pie

ggsave("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/Categories_Terr.png", width = 6,   height = 4.25, dpi=1100)
```
</div>

<div class = "col-md-5">
``` {r echo=FALSE}
funcTerr$Funding <- signif(funcTerr$Funding, 5)
funcTerr$Funding <- formattable::currency(funcTerr$Funding, digits = 0L)
terrFuncTable <-funcTerr%>%select(Category, Funding)%>%kbl(row.names=FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, html_font="Open Sans", font_size = 14, position="center", row_label_position = r)%>%
  add_header_above(c("Territory Funding by Category"=2))%>%
  row_spec(1:5, color  = "black")%>%
  column_spec(1:2, "4cm")
terrFuncTable
save_kable(terrFuncTable,"C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/TerrFunctable.png")
```
</div>
### Funding by Award Type
<div class = "row">
<div class = "col-md-7" style="padding-right:25px;">
``` {r echo=FALSE}
fundByType_Terr <- funData %>% dplyr::filter(recipient_state %in% targetN)%>%
  group_by(award_type) %>% 
  dplyr::summarise(
  Funding = sum(transaction_obligated_amount)
  )%>% as.data.frame()

fundByType_Terr$Percent <- 100*(fundByType_Terr$Funding/sum(fundByType_Terr$Funding))
fundByType_Terr$Percent <- format(round(fundByType_Terr$Percent, 2), nsmall=2)
fundByType_Terr$Percent <- as.numeric(fundByType_Terr$Percent)

fundByType_Terr$zone.start <- with(fundByType_Terr, c(0, cumsum(Percent)[-length(Percent)]))  
fundByType_Terr$zone.end <- with(fundByType_Terr, cumsum(Percent))  
fundByType_Terr$label.point <- with(fundByType_Terr, (100-(zone.start + zone.end) / 2)) 

colnames(fundByType_Terr)[1] <- "Type"
fundByType_Terr$Type <- str_to_title(fundByType_Terr$Type)
fundByType_Terr$Type[4] <- "Direct Payment Specified Use"
fundByType_Terr$Type[6] <- "Other Financial Assistance"
fundByType_Terr$Type[1] <- "Block Grant"
fundByType_Terr$Type[2] <- "Cooperative Agreement"
fundByType_Terr$Type[5] <- "Formula Grant"
fundByType_Terr$Type[7] <- "Project Grant"

pie <- ggplot(fundByType_Terr, aes(x="", y=Percent, fill=Type))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=c("#900C3F", "#D4442D", "#FCB616", "#75C7D3", "#00869F", "#C799C8", "#D2D2D2", "#585757")) +
  geom_text_repel(data = fundByType_Terr,
                   aes(y = label.point, label = ifelse(Percent>5,paste0(Percent, "%"),""),family="Open Sans"),
                   size = 4.5, nudge_x = 1, show.legend = FALSE, max.overlaps =20) +
  ggtitle("U.S. Territories Funding by Award Type")+
  theme_void(base_family= "Open Sans")+ 
  theme(plot.title = element_text(size=12))
pie

ggsave("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/award_type.png", width = 6,   height = 4, dpi=1100)
```
</div>

<div class = "col-md-5">
``` {r echo=FALSE}
fundByType_Terr$Funding <- signif(fundByType_Terr$Funding, 5)
fundByType_Terr$Funding <- formattable::currency(fundByType_Terr$Funding, digits = 0L)
terrTypeTable <-fundByType_Terr%>%select(Type, Funding)%>%kbl(row.names=FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, html_font="Open Sans", font_size = 14, position="center", row_label_position = r)%>%
  add_header_above(c("Territory Funding by Type"=2))%>%
  row_spec(1:5, color  = "black")%>%
  column_spec(1:2, "4cm")
terrTypeTable
save_kable(terrTypeTable,"C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/terrTypeTable.png")
```
</div>
### Energy Funding
<div class = "row">
<div class = "col-md-6" style="padding-right:20px;">
``` {r echo=FALSE}
###Weatherization Assistance Program Funding TERRITORIES
WAPdata_terr <- read.csv("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/EnergyFunding_Territories/EnergyFunding_Territories.csv")%>%select(obligated_amount_from_IIJA_supplemental, program_activities_funding_this_award, recipient_name, cfda_numbers_and_titles, recipient_state_name)%>%filter(cfda_numbers_and_titles == "81.042: WEATHERIZATION ASSISTANCE FOR LOW-INCOME PERSONS")

colnames(WAPdata_terr)[1] <- "Funding"
colnames(WAPdata_terr)[3] <- "Recipient"
WAPdata_terr$Recipient <- paste(str_to_title(WAPdata_terr$Recipient))
WAPdata_terr$Funding <- signif(WAPdata_terr$Funding, 5)
WAPdata_terr$Funding <- currency(WAPdata_terr$Funding, digits = 0L)
WAPdata_terr <- WAPdata_terr[order(WAPdata_terr$Funding,decreasing=TRUE),]
WAPdata_terr$Recipient[1] <- "Departamento De Desarrollo Economico Y Comercio (Puerto Rico)"
WAPtable<- WAPdata_terr%>% select(Recipient, Funding)%>% kbl(row.names=FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, html_font="Open Sans", font_size = 14)%>%  
  add_header_above(c("Weatherization Assistance Program Funding"=2))%>%
  row_spec(1:5, color  = "black")%>%
  column_spec(1, "12cm")%>%
  column_spec(2, "3cm")
WAPtable
save_kable(WAPtable,"C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/WAPtable_terr.png")
WAPtotal_terr <- sum(WAPdata_terr$Funding)
```
</div>

<div class = "col-md-6">
```{r, echo=FALSE, fig.align='center'}
###State Energy Office Funding TERRITORIES
SEOdata_terr <- read.csv("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/EnergyFunding_Territories/EnergyFunding_Territories.csv")%>%select(obligated_amount_from_IIJA_supplemental, program_activities_funding_this_award, recipient_name, cfda_numbers_and_titles, recipient_state_name)%>%filter(cfda_numbers_and_titles == "81.041: STATE ENERGY PROGRAM")

colnames(SEOdata_terr)[1] <- "Funding"
colnames(SEOdata_terr)[3] <- "Recipient"
SEOdata_terr$Recipient <- paste(str_to_title(SEOdata_terr$Recipient))
SEOdata_terr$Funding <- signif(SEOdata_terr$Funding, 5)
SEOdata_terr$Funding <- currency(SEOdata_terr$Funding, digits = 0L)
SEOdata_terr <- SEOdata_terr[order(SEOdata_terr$Funding,decreasing=TRUE),]
SEOdata_terr$Recipient[5] <- "Departamento De Desarrollo Economico Y Comercio (Puerto Rico)"

SEOtable<-SEOdata_terr%>% select(Recipient, Funding)%>% kbl(row.names=FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), html_font="Open Sans", font_size = 14, full_width = F)%>%
  add_header_above(c("State Energy Program Funding"=2))%>%
  row_spec(1:5, color  = "black")%>%
  column_spec(1, "12cm")%>%
  column_spec(2, "3cm")
SEOtable
save_kable(SEOtable,"C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/SEOtable_terr.png")
SEOtotal_terr <- sum(SEOdata_terr$Funding)

# #function and award type data, filter to Territories minus PR
# funcTerr <-dplyr::filter(funData, recipient_state=="AS"|recipient_state=="MP"|recipient_state=="GU"|recipient_state=="VI")%>%
#   group_by(budget_subfunction) %>% dplyr::summarise(
#   Funding = sum(transaction_obligated_amount)
#   )%>% as.data.frame()
# 
# funcTerr$Percent <- 100*(funcTerr$Funding/sum(funcTerr$Funding))
# funcTerr$Percent <- format(round(funcTerr$Percent, 2), nsmall=2)
# funcTerr$Percent <- as.numeric(funcTerr$Percent)
# 
# funcTerr$zone.start <- with(funcTerr, c(0, cumsum(Percent)[-length(Percent)]))  
# funcTerr$zone.end <- with(funcTerr, cumsum(Percent))  
# funcTerr$label.point <- with(funcTerr, (100-(zone.start + zone.end) / 2)) 
# 
# colnames(funcTerr)[1] <- "Category"
# 
# pie <- ggplot(funcTerr, aes(x="", y=Percent, fill=Category))+
#   geom_bar(width = 1, stat = "identity")+
#   coord_polar("y", start=0)+
#   scale_fill_manual(values=c("#FFA07A","#900C3F","#C70039", "#D4442D", "#00869F", "#6171AF","#DCB6D0","#C799C8","#7B7B7B","#585757")) +
#   geom_text_repel(data = funcTerr,
#                    aes(y = label.point, label = ifelse(Percent>10,paste0(Percent, "%"),""), family="Open Sans"),
#                    size = 4.5, nudge_x = 1, show.legend = FALSE)+
#   ggtitle("American Samoa, Guam, Northern Mariana Islands, and Virgin Islands")+
#   theme_void(base_family = "Open Sans")
# pie
# ggsave("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/FuncTerr_cat.png", width = 6,   height = 4.25, dpi=1100)

```
</div>
### American Samoa
<div class = "row">
<div class = "col-md-7" style="padding-right:25px;">
```{r, echo=FALSE, fig.align='center'}
#function and award type data, filter to American Samoa
funcAS <-dplyr::filter(funData, recipient_state=="AS")%>%
  group_by(budget_subfunction) %>% dplyr::summarise(
  Funding = sum(transaction_obligated_amount)
  )%>% as.data.frame()

funcAS$Percent <- 100*(funcAS$Funding/sum(funcAS$Funding))
funcAS$Percent <- format(round(funcAS$Percent, 2), nsmall=2)
funcAS$Percent <- as.numeric(funcAS$Percent)

funcAS$zone.start <- with(funcAS, c(0, cumsum(Percent)[-length(Percent)]))  
funcAS$zone.end <- with(funcAS, cumsum(Percent))  
funcAS$label.point <- with(funcAS, (100-(zone.start + zone.end) / 2)) 

colnames(funcAS)[1] <- "Category"

pie <- ggplot(funcAS, aes(x="", y=Percent, fill=Category))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=c("#900C3F","#C70039", "#D4442D", "#00869F", "#6171AF","#DCB6D0","#C799C8","#585757")) +
  geom_text_repel(data = funcAS,
                   aes(y = label.point, label = ifelse(Percent>10,paste0(Percent, "%"),""), family="Open Sans"),
                   size = 4.5, nudge_x = 1, show.legend = FALSE)+
  ggtitle("American Samoa Funding")+
  theme_void(base_family = "Open Sans")
pie
ggsave("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/AmericanSam_cat.png", width = 6,   height = 4.25, dpi=1100)

```
</div>

<div class = "col-md-5">
``` {r echo=FALSE}
funcAS$Funding <- signif(funcAS$Funding, 5)
funcAS$Funding <- currency(funcAS$Funding, digits = 0L)
funcAS%>% select(Category, Funding)%>% kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, html_font="Open Sans", font_size = 14)%>%  
  add_header_above(c("American Samoa Funding by Category"=2))%>%
  row_spec(1:8, color  = "black")%>%
  column_spec(1, "10cm")%>%
  column_spec(2, "3cm")
```
</div>
</div>
<div class = "row">
<br>
<br>
<br>
</div>
#### Spotlight on Projects
American Samoa Government Department of Commerce: The National Oceanic and Atmospheric Administration committed $203,000 to American Samoa for wetland delineation, wetland monitoring, training, and community outreach, as well as to create recommendations for permitting policies and nature-based restoration projects. 

### Guam
<div class = "row">
<div class = "col-md-7" style="padding-right:25px;">
```{r, echo=FALSE, fig.align='center'}
#function and award type data, filter to GUAM
funcGU <-dplyr::filter(funData, recipient_state=="GU")%>%
  group_by(budget_subfunction) %>% dplyr::summarise(
  Funding = sum(transaction_obligated_amount)
  )%>% as.data.frame()

funcGU$Percent <- 100*(funcGU$Funding/sum(funcGU$Funding))
funcGU$Percent <- format(round(funcGU$Percent, 2), nsmall=2)
funcGU$Percent <- as.numeric(funcGU$Percent)

funcGU$zone.start <- with(funcGU, c(0, cumsum(Percent)[-length(Percent)]))  
funcGU$zone.end <- with(funcGU, cumsum(Percent))  
funcGU$label.point <- with(funcGU, (100-(zone.start + zone.end) / 2)) 

colnames(funcGU)[1] <- "Category"

pie <- ggplot(funcGU, aes(x="", y=Percent, fill=Category))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=c("#900C3F","#C70039", "#D4442D", "#00869F", "#6171AF","#DCB6D0","#C799C8","#585757")) +
  geom_text_repel(data = funcGU,
                   aes(y = label.point, label = ifelse(Percent>10,paste0(Percent, "%"),""), family="Open Sans"),
                   size = 4.5, nudge_x = 1, show.legend = FALSE)+
  ggtitle("Guam Funding")+
  theme_void(base_family = "Open Sans")
pie

ggsave("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/Guam_cat.png", width = 6,   height = 4.25, dpi=1100)
```
</div>
<div class = "col-md-5">

``` {r echo=FALSE}

funcGU$Funding <- signif(funcGU$Funding, 5)
funcGU$Funding <- currency(funcGU$Funding, digits = 0L)
funcGU%>% select(Category, Funding)%>% kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, html_font="Open Sans", font_size = 14)%>%  
  add_header_above(c("Guam Funding by Category"=2))%>%
  row_spec(1:8, color  = "black")%>%
  column_spec(1, "10cm")%>%
  column_spec(2, "3cm")
```
</div>
</div>
<div class = "row">
<br>
<br>
<br>
</div>
#### Spotlight on Projects
Teleguam Holdings: The Department of Agriculture committed $29 million to Teleguam Holdings to provide internet connection to 8,000 households in Guam.

### Northern Mariana Islands
<div class = "row">
<div class = "col-md-7" style="padding-right:25px;">
```{r, echo=FALSE, fig.align='center'}
#function and award type data, filter to Northern Mariana Islands
funcMP <-dplyr::filter(funData, recipient_state=="MP")%>%
  group_by(budget_subfunction) %>% dplyr::summarise(
  Funding = sum(transaction_obligated_amount)
  )%>% as.data.frame()

funcMP$Percent <- 100*(funcMP$Funding/sum(funcMP$Funding))
funcMP$Percent <- format(round(funcMP$Percent, 2), nsmall=2)
funcMP$Percent <- as.numeric(funcMP$Percent)

funcMP$zone.start <- with(funcMP, c(0, cumsum(Percent)[-length(Percent)]))  
funcMP$zone.end <- with(funcMP, cumsum(Percent))  
funcMP$label.point <- with(funcMP, (100-(zone.start + zone.end) / 2)) 

colnames(funcMP)[1] <- "Category"

pie <- ggplot(funcMP, aes(x="", y=Percent, fill=Category))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=c("#900C3F","#C70039", "#D4442D", "#00869F", "#6171AF","#DCB6D0","#C799C8","#585757")) +
  geom_text_repel(data = funcMP,
                   aes(y = label.point, label = ifelse(Percent>10,paste0(Percent, "%"),""), family="Open Sans"),
                   size = 4.5, nudge_x = 1, show.legend = FALSE)+
  ggtitle("Northern Mariana Islands Funding")+
  theme_void(base_family = "Open Sans")
pie

ggsave("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/Mariana_cat.png", width = 6,   height = 4.25, dpi=1100)

```
</div>
<div class = "col-md-5">
``` {r echo=FALSE}
funcMP$Funding <- signif(funcMP$Funding, 5)
funcMP$Funding <- currency(funcMP$Funding, digits = 0L)
funcMP%>% select(Category, Funding)%>% kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, html_font="Open Sans", font_size = 14)%>%  
  add_header_above(c("Northern Mariana Islands Funding by Category"=2))%>%
  row_spec(1:8, color  = "black")%>%
  column_spec(1, "10cm")%>%
  column_spec(2, "3cm")
```
</div>
</div>
<div class = "row">
<br>
<br>
<br>
</div>

#### Spotlight on Projects 
Commonwealth Utilities Corporation: The Environmental Protection Agency committed $20.6 million to improve the drinking water system on the Northern Mariana Islands, including by replacing and upgrading water mains and pressurizing the transmission and distribution system. These projects will aid in providing reliable, potable drinking water to every customer.

### Puerto Rico
<div class = "row">
<div class = "col-md-7" style="padding-right:25px;">
```{r, echo=FALSE, fig.align='center'}
#function and award type data, filter to Puerto Rico
funcPR <-dplyr::filter(funData, recipient_state=="PR")%>%
  group_by(budget_subfunction) %>% dplyr::summarise(
  Funding = sum(transaction_obligated_amount)
  )%>% as.data.frame()

funcPR$Percent <- 100*(funcPR$Funding/sum(funcPR$Funding))
funcPR$Percent <- format(round(funcPR$Percent, 2), nsmall=2)
funcPR$Percent <- as.numeric(funcPR$Percent)

funcPR$zone.start <- with(funcPR, c(0, cumsum(Percent)[-length(Percent)]))  
funcPR$zone.end <- with(funcPR, cumsum(Percent))  
funcPR$label.point <- with(funcPR, (100-(zone.start + zone.end) / 2)) 

colnames(funcPR)[1] <- "Category"

pie <- ggplot(funcPR, aes(x="", y=Percent, fill=Category))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=c("#FFA07A","#900C3F", "#C70039", "#D4442D", "#00869F", "#6171AF","#DCB6D0","#C799C8", "#7B7B7B","#585757","#17202A")) +
  geom_text_repel(data = funcPR,
                   aes(y = label.point, label = ifelse(Percent>10,paste0(Percent, "%"),""), family="Open Sans"),
                   size = 4.5, nudge_x = 1, show.legend = FALSE)+
  ggtitle("Puerto Rico")+
  theme_void(base_family = "Open Sans")
pie

ggsave("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/PuertoRico_cat.png", width = 6,   height = 4.25, dpi=1100)

```
</div>
<div class = "col-md-5">

``` {r echo=FALSE}
funcPR$Funding <- signif(funcPR$Funding, 5)
funcPR$Funding <- currency(funcPR$Funding, digits = 0L)
funcPR%>% select(Category, Funding)%>% kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, html_font="Open Sans", font_size = 14)%>%  
  add_header_above(c("Puerto Rico Funding by Category"=2))%>%
  row_spec(1:8, color  = "black")%>%
  column_spec(1, "10cm")%>%
  column_spec(2, "3cm")
```
</div>
#### Spotlight on Projects
Institute for Socio-Ecological Research: The Department of Commerce will commit $10.6 million to construct five acres of coral reef at three locations in Puerto Rico. The project will reintroduce slow-growing, reef-building coral species, including threatened species. 

### Virgin Islands
<div class = "row">
<div class = "col-md-7" style="padding-right:25px;">
```{r, echo=FALSE, fig.align='center'}
#function and award type data, filter to Virgin Islands
funcVI <-dplyr::filter(funData, recipient_state=="VI")%>%
  group_by(budget_subfunction) %>% dplyr::summarise(
  Funding = sum(transaction_obligated_amount)
  )%>% as.data.frame()

funcVI$Percent <- 100*(funcVI$Funding/sum(funcVI$Funding))
funcVI$Percent <- format(round(funcVI$Percent, 2), nsmall=2)
funcVI$Percent <- as.numeric(funcVI$Percent)

funcVI$zone.start <- with(funcVI, c(0, cumsum(Percent)[-length(Percent)]))  
funcVI$zone.end <- with(funcVI, cumsum(Percent))  
funcVI$label.point <- with(funcVI, (100-(zone.start + zone.end) / 2)) 

colnames(funcVI)[1] <- "Category"

pie <- ggplot(funcVI, aes(x="", y=Percent, fill=Category))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=c("#FFA07A","#900C3F", "#C70039", "#00869F", "#6171AF","#DCB6D0","#C799C8", "#7B7B7B","#585757")) +
  geom_text_repel(data = funcVI,
                   aes(y = label.point, label = ifelse(Percent>10,paste0(Percent, "%"),""), family="Open Sans"),
                   size = 4.5, nudge_x = 1, show.legend = FALSE)+
  ggtitle("Virgin Islands")+
  theme_void(base_family = "Open Sans")
pie

ggsave("C:/Users/GraceParker/Downloads/2023_FederalFundingTracker/VirginIs_cat.png", width = 6,   height = 4.25, dpi=1100)

```

</div>
<div class = "col-md-5">
``` {r echo=FALSE}
funcVI$Funding <- signif(funcVI$Funding, 5)
funcVI$Funding <- currency(funcVI$Funding, digits = 0L)
funcVI%>% select(Category, Funding)%>% kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, html_font="Open Sans", font_size = 14)%>%  
  add_header_above(c("U.S. Virgin Islands Funding by Category"=2))%>%
  row_spec(1:8, color  = "black")%>%
  column_spec(1, "10cm")%>%
  column_spec(2, "3cm")
```
</div>
<div class = "row">
<br>
<br>
<br>
</div>

#### Spotlight on Projects
Island Conservation: The Department of Interior committed $250,000 to Island Conservation to aid in invasive species eradication and create a refuge for the endangered Virgin Islands tree boa.